Linear analysis of the vertical shear instability in astrophysical disks

### 2015-02-03 numerical.tex
completed rough summary
updating

\section{Numerical solutions}
In this section we numerically solve the linear stability problem
corresponding to Eq. \ref{lin_mass}---\ref{lin_energy}. 
We relax the thin-disk, low-frequency and nearly-Keplerian
approximations used in the discussion above, and 
use the full, non-Keplerian expressions of $\rho$, $\Omega$ and
$\kappa$, so that we may consider vertical domains up to $z\sim r$.

We solve the linear problem by expanding the
perturbations in Chebyshev polynomials $T_l$ up to $l=512$
and discretizing the equations on a grid with
$z\in[-\zmax,\zmax]$. At the vertical boundaries we impose a free
surface, 
\begin{align}
  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0 \quad \text{at } z=\pm\zmax,
\end{align}
where $\bm{\xi}$ is the Lagrangian displacement. This boundary
condition was found to be most convenient for numerical
implementation. The above discretization procedure converts the linear  
system of differential equations to a standard matrix eigenvalue
problem.

Unless otherwise stated, our disk model is a nearly-vertically
isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
$\epsilon=0.1$.

@@

\subsection{VSI in a neutrally stratified disk}\label{vertiso_pertiso} 
We first calculate the VSI in a neutrally straitified disk ($N_z^2\equiv0$) by setting
$\gamma=\Gamma$, and effectively switch off thermal relaxation by
choosing $\beta=10^9$. We set the vertical domain to $\zmax =
10\epsilon r$.  Here, we consider a perturbation wavenumber
$\khat=20\pi$ or $k_x = 200\pi/r$. This setup is very similar to that
adopted in \cite{mcnally14}, who calculated the VSI in vertically
isothermal disks subject to isothermal perturbations
($\Gamma=\gamma=1$).  This allows us to test our numerical code.

Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
\ii\nu$. The fundamental VSI mode, with one node in $W$, has
the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}. 
The eigenfunction $W$ is qualitatively proportional to 
$z$ (especially for $z\to0$),  as expected for the fundamental mode
according to explicit solutions developed in \S\ref{iso_poly}.    
These plots are consistent with those in \cite{mcnally14}.

The expected growth rate of the fundamental VSI, calculated from
Eq. \ref{simple_growth} with $L=1$, and the numerically-obtained value
are  
\begin{align*}
  \nu = 0.2606\epsilon\Omega_k &\quad \text{(from
    Eq. \ref{simple_growth})},\\
  \nu = 0.2577 \epsilon \Omega_k &\quad \text{(numerical)}.
\end{align*}
This agreement is surpsingly good, given that
Eq. \ref{simple_growth} assumes a thin, nearly-Keplerian disk in the 
low-frequency approximation, and
imposes a different boundary condition to that in the numerical
calculation. The eigenfrequency is not very sensitive to the vertical boundary
condition because the expression for $\sigma$, given by
Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the 
weight function, which rapidly decays away from the midplane.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
  \caption{Eigenvalues of the VSI in a neutrally-stratified,
    nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$) with 
    $(p,q,\epsilon)=(-1.5,-1,0.1)$, and the perturbation radial
    wavenumber is $\khat=20\pi$. \label{lowfreq_eigen}
  }
\end{figure}

@@

\subsection{VSI in a stably stratified disk}
We demonstrate the (strong) stabilizing effect of a  
positive vertical entropy gradient ($N_z^2\geq0$) by setting $\gamma=2$.  
Since we expect the perturbations to decay rapidly
away from the midplane (see \S\ref{analytic_adia}), we consider a
smaller domain with $\zmax = 2\epsilon r$. 
Other disk and perturbation parameters are the same as in
\S\ref{vertiso_pertiso}.

Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
problem. The growth rates are much smaller than
those in the previous section. The fundamental VSI mode is that with
$\mathrm{max}\nu$. Its expected and numerically-calculated growth
rates compares well, with 
\begin{align*}
  \nu = 0.03751 \epsilon \Omega_k &\quad \text{(from
    Eq. \ref{gam2_growth_rate})},\\
  \nu = 0.03672 \epsilon \Omega_k &\quad \text{(numerical)}.
\end{align*}
The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
and clearly show that perturbations are rapidly stabilized away from
the midplane. For the adopted disk and perturbation parameters,
Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$, corresponding
to a characteristic decay length scale of $\sqrt{2/|\alpha|}\epsilon r\simeq 0.2\epsilon r$,
which is consistent with that observed in  Fig. \ref{lowfreq_eigenfunc_adia}.

This decay lengthscale is rougly the height within which
vertical shear is larger than the buoyancy frequency squared. That is,
\begin{align*}
\left|r\frac{d\Omega^2}{dz}\right|\gtrsim N_z^2
\end{align*}
for
\begin{align*}
  \left|\frac{z}{\epsilon r}\right| \lesssim \frac{\epsilon|q|\gamma}{\gamma-1},
\end{align*}
in a thin, vertically isothermal disk. This evaluates to $|z|\lesssim
0.2\epsilon r$ for the current disk model, consistent with
Fig. \ref{lowfreq_eigenfunc_adia}.

Furthermore, the growth rate to should be no larger than the
maximum vertical shear available within this decay lengthscale. Using 
Eq. \ref{max_growth} as a rough guide, 
\begin{align}
  \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2\gamma}{2(\gamma-1)},
\end{align}
or $\nu< 0.1\epsilon\Omega_k$ in our disk model,
consistent with numerical results.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
  \caption{Eigenvalues in for the
    VSI in a nearly vertically isothermal disk
    ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
    evolved adiabatically with $\gamma=2$. The perturbation radial
    wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
  }
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
  \caption{Fundamental VSI mode in a nearly vertically
    isothermal disk ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
    evolved adiabatically
    with $\gamma=2$. This eigenfunction 
    corresponds to the top-left eigenvalue displayed in 
    Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
    The  real (imaginary) part of $\delta v_z$ are shown as solid
    (dotted) lines. 
    \label{lowfreq_eigenfunc_adia}
  }
\end{figure}

Although a positive vertical entropy gradient is stabilizing, we note
that $N_z^2\propto z^2$  while vertical shear $rd\Omega^2/dz\propto
z$, so that there is always a region sufficiently close to the
midplane in which vertical shear dominates over bouyancy and the VSI
can operate. However, the growth rates are small and the instability
only affects the regions close the midplane. These modes may not be
important in practice.

\subsection{Other adiabatic indices}
To see the transition from neutrally-bouyant to stably stratified cases above, 
we repeated the previous calculation with $\gamma\in[1,2]$. 
We restore the vertical domain to $\zmax=10\epsilon r$. 
Fig. \ref{adia_growth_num} shows the eigenfrequencies of the
fundamental VSI mode.

%To select the appropriate
%eigenvalues we vary the parameters slowly away from a reference solution.   
% The numerically-determined eigenfrequencies are qualitatively
% consistent with the discussion in \S\ref{adia_compare_gam1}.
There is good agreement between Fig. \ref{adia_growth_num} and
Fig. \ref{adia_growth} for $\khat=k_xH_\mathrm{iso}\gtrsim10$. There is a
mismatch for smaller wavenumbers, especially in the growth
rates. Actual growth rates decrease slower as $\khat\to0$
than that in Fig. \ref{adia_growth}. Disagreement at small
$\khat$ is not unexpected because the low-frequency
approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
$|\omega|\sim\Omega_k$ for small $\khat$.

Given the number of approximations made in \S\ref{adia_compare_gam1},
Fig. \ref{adia_growth_num} compares well with 
Fig. \ref{adia_growth}. The main qualitative features of the full
numerical solution in Fig. \ref{adia_growth_num} are captured by
Fig. \ref{adia_growth}: stabilization for 
$\khat\to0$ and $\khat\to\infty$,
stabilization for increasing $\gamma$ (but ineffective at small $\khat$), 
and the shift of the most unstable wavenumber to smaller values as
$\gamma$ is increased.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_real1}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag1}
  \caption{Real frequency (top) and growth rate 
    (bottom) of the fundamental VSI in a nearly vertically   
    isothermal disks ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
    subject to adiabatic perturbations with different values of $\gamma$. 
    The eigenfrequencies are calculated by numerically
    solving the full linear eigenvalue problem. This plot is to
    be compared with Fig. \ref{adia_growth}. \label{adia_growth_num}}   
\end{figure}

\subsection{Effect of thermal relaxation}
We now introduce thermal relaxation. Here we consider a slightly
different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
adopted in simulations performed by \cite{nelson13}. 
Fig. \ref{relax_growth_num} plots the
growth rates of the fundamental mode function of $\khat$.

%corrugation mode
% We first obtain the eigenvalues for the
% fundamental mode for $k_xH_\mathrm{iso}=0$, then vary parameters
% slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.

Fig. \ref{relax_growth_num} is largely consistent with 
Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
where the low-frequency approximation fails. 
For $\beta=0.063$ at $\khat\gtrsim50$ we found eigenfunctions had 
large perturbations near the vertical boundaries, which may be 
responsible for the non-smoothness in the curve at large $\khat$.

%likely attributable to the free surface boundary condition imposed in
%the numerical calculation (cf. regularity condition at infinity in
%\S\ref{disp_relax}).

@@

\begin{figure}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
   0cm]{figures/compare_eigen_imag_bloop} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/compare_eigen_imag_bloop2} 
  \caption{Growth rates of the fundamental VSI mode in 
    nearly-vertically isothermal disks ($\Gamma=1.011$) with
    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
    range of thermal relaxation timescales   
    $\beta$. The eigenfrequencies are calculated by numerically 
    solving the full linear eigenvalue problem. \label{relax_growth_num}}   
\end{figure}

@@

%\citeauthor{nelson13} find a thermal relaxation timescale of
% $\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
% hydrodynamic simulations. This is consistent with our numerical
% results.

@@

>>>>

\section{Numerical solutions}
In this section we numerically solve the linear stability problem
corresponding to Eq. \ref{lin_mass}---\ref{lin_energy}. 
We relax the thin-disk, low-frequency and nearly-Keplerian
approximations used in the discussion above, and 
use the full, non-Keplerian expressions of $\rho$, $\Omega$ and
$\kappa$.
%, so that we may consistently consider vertical domains up to
%$z\sim r$.

We solve the linear problem by expanding the
perturbations in Chebyshev polynomials $T_l$ up to $l=512$
and discretizing the equations on a grid with
$z\in[-\zmax,\zmax]$. At the vertical boundaries we impose a free
surface, 
\begin{align}
  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0 \quad \text{at } z=\pm\zmax,
\end{align}
where $\bm{\xi}$ is the Lagrangian displacement. This boundary
condition is convenient for numerical implementation since it conforms
to a standard eigenvalue problem. The above discretization procedure
converts the linear system of differential equations to a set of 
algebraic equations, for which we use matrix routines in the LAPACK
package to solve.

@@

\subsection{VSI in a locally isothermal, neutrally stratified disk}\label{vertiso_pertiso} 
%  disk model is a nearly-vertically
% isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
% $\epsilon=0.1$. 
We first calculate the VSI in a nearly vertcally isothermal, neutrally
straitified disk ($N_z^2\equiv0$) by setting
$\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk parameters are 
$(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain is $\zmax =
10\epsilon r$. This setup is very similar to that adopted in
\cite{mcnally14}, who calculated the VSI in vertically isothermal
disks subject to isothermal perturbations ($\Gamma=\gamma=1$).  This
allows us to test our numerical code.

%behaviour as func of kx

Fig. \ref{iso_eigen_kx} compares the numerically-obtained eigenvalues
for the fundamental mode and that caluclated from
Eq. \ref{simple_growth} (with $L=1$). For completeness we
included cases with $\khat\ll 1$ even though the local
approximation does not hold in this regime. The agreement is good for
$\khat\gtrsim10$. The mismatch is significant for $\khat\ll1$ since
the low-frequency approximation used to obtain Eq. \ref{simple_growth}
is no longer valid: $|\omega|\sim\Omega_k$ as $\khat\to 0$. This means
that our analytic results should only be applied to perturbations with
$\khat\gg 1$, i.e. small radial-lengthscales.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_eigen_imag_iso} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/compare_eigen_real_iso}
  \caption{Growth rate (top) and real frequency (bottom) of the
    fundamental VSI mode in a disk with $\gamma=\Gamma=1.011$,
    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
    (analytically).  
    \label{iso_eigen_kx} 
  }
\end{figure}

%particular kx 
As an example of an unstable mode, we consider a perturbation
wavenumber $\khat=20\pi$ \citep[or $k_x = 200\pi/r$ as used
in][]{mcnally14}. Fig. \ref{lowfreq_eigen} shows the eigenvalues
$\sigma = \omega + \ii\nu$. The fundamental VSI mode, with one node in $W$, has
the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}. 
The eigenfunction $W$ is qualitatively proportional to 
$z$ (especially for $z\to0$),  as expected for the fundamental mode
according to explicit solutions developed in \S\ref{iso_poly}.    
These plots are consistent with those in \cite{mcnally14}.

The growth rate of the fundamental VSI calculated from
Eq. \ref{simple_growth} with $L=1$ and the numerically-obtained value
are  
\begin{align*}
  \nu = 0.2606\epsilon\Omega_k &\quad \text{(from
    Eq. \ref{simple_growth})},\\
  \nu = 0.2575 \epsilon \Omega_k &\quad \text{(numerical)}.
\end{align*}
This agreement is surpsingly good, given that
Eq. \ref{simple_growth} assumes a thin, nearly-Keplerian disk in the 
low-frequency approximation, and 
imposes a different boundary condition to that in the numerical
calculation. These values are also close to that found by
\cite{mcnally14}. The eigenfrequency is not sensitive to the
vertical boundary condition because the expression for $\sigma$, given
by Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the  
weight function, which rapidly decays away from the midplane.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
  \caption{Eigenvalues of the VSI in a neutrally-stratified,
    nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$) with 
    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and $\beta=10^{-3}$. The
    perturbation wavenumber is $\khat=20\pi$. \label{lowfreq_eigen} 
  }
\end{figure}

@@

\subsection{Effect of thermal relaxation}
We now consider stably stratified disks ($\gamma > \Gamma$) with
thermal relaxation $\beta\in[10^{-3},1]$. Here we use a slightly 
different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
adopted in simulations performed by \cite{nelson13}. Other parameters
are unchanged from the previous section.

Fig. \ref{relax_growth_num} shows the growth rates of the fundamental  
mode as a function of $\khat$. This plot is largely consistent with  
Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
where the low-frequency approximation fails. (The local model not valid
for $\khat\ll1$ anyway). For $\khat\gtrsim 10$ introducing thermal
relaxation rapidly stabilizes the fundamental mode.

% For $\beta=0.063$ at $\khat\gtrsim50$ we found eigenfunctions had 
% large perturbations near the vertical boundaries, which may be 
% responsible for the non-smoothness in the curve at large $\khat$. 
%corrugation mode
% We first obtain the eigenvalues for the
% fundamental mode for $k_xH_\mathrm{iso}=0$, then vary parameters
% slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.  
%likely attributable to the free surface boundary condition imposed in
%the numerical calculation (cf. regularity condition at infinity in
%\S\ref{disp_relax}).

@@

\begin{figure}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
   0cm]{figures/compare_eigen_imag_bloop} 
  % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  % 1cm]{figures/compare_eigen_imag_bloop2} 
  \caption{Growth rates of the fundamental VSI mode in 
    nearly-vertically isothermal disks ($\Gamma=1.011$) with
    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
    range of thermal relaxation timescales   
    $\beta$. The eigenfrequencies are calculated by numerically 
    solving the full linear eigenvalue problem. \label{relax_growth_num}}   
\end{figure}

@@

For this disk model, \citeauthor{nelson13} found a thermal relaxation
timescale of $\beta\gtrsim 0.6$ stabilized the VSI in non-linear
hydrodynamic simulations. This is consistent with our numerical and
analytic results.

\subsubsection{Dependence on disk parameters}
We test our estimate for the critical relaxation time
$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}). We use the above disk
parameters as a fiducial case and vary $q\in[-0.2,-1.2]$,
$\gamma\in[1.2,2.0]$ and $\epsilon\in[]$
separately. The numerically-obtained $\beta_\mathrm{crit}$ is shown in
Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond}.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
  0cm]{figures/bcrit_compare_q.ps} 
  \caption{Dependence of the upper limit to the thermal relaxation timescale
    $\beta_\mathrm{crit}$ for the fundamental VSI mode on disk
    parameters. The fiducial parameters are $\Gamma=1.011$,
    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. Top: varying
    $q\in[-1.2,-0.2]$; middle: varying $\gamma\in[1.2,2.0]$; bottom:
    varying $\epsilon\in[]$. The perturbation wavenumber is
    $\khat=20$. {\bf waiting for high resolution}
    \label{bcrit_compare}}  
\end{figure}

@@

% \subsection{VSI in a stably stratified disk}
% We demonstrate the (strong) stabilizing effect of a  
% positive vertical entropy gradient ($N_z^2\geq0$) by setting $\gamma=2$.  
% Since we expect the perturbations to decay rapidly
% away from the midplane (see \S\ref{analytic_adia}), we consider a
% smaller domain with $\zmax = 2\epsilon r$. 
% Other disk and perturbation parameters are the same as in
% \S\ref{vertiso_pertiso}.

% Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
% problem. The growth rates are much smaller than
% those in the previous section. The fundamental VSI mode is that with
% $\mathrm{max}\nu$. Its expected and numerically-calculated growth
% rates compares well, with 
% \begin{align*}
%   \nu = 0.03751 \epsilon \Omega_k &\quad \text{(from
%     Eq. \ref{gam2_growth_rate})},\\
%   \nu = 0.03672 \epsilon \Omega_k &\quad \text{(numerical)}.
% \end{align*}
% The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
% and clearly show that perturbations are rapidly stabilized away from
% the midplane. For the adopted disk and perturbation parameters,
% Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$, corresponding
% to a characteristic decay length scale of $\sqrt{2/|\alpha|}\epsilon r\simeq 0.2\epsilon r$,
% which is consistent with that observed in  Fig. \ref{lowfreq_eigenfunc_adia}.

% This decay lengthscale is rougly the height within which
% vertical shear is larger than the buoyancy frequency squared. That is,
% \begin{align*}
% \left|r\frac{d\Omega^2}{dz}\right|\gtrsim N_z^2
% \end{align*}
% for
% \begin{align*}
%   \left|\frac{z}{\epsilon r}\right| \lesssim \frac{\epsilon|q|\gamma}{\gamma-1},
% \end{align*}
% in a thin, vertically isothermal disk. This evaluates to $|z|\lesssim
% 0.2\epsilon r$ for the current disk model, consistent with
% Fig. \ref{lowfreq_eigenfunc_adia}.

% Furthermore, the growth rate to should be no larger than the
% maximum vertical shear available within this decay lengthscale. Using 
% Eq. \ref{max_growth} as a rough guide, 
% \begin{align}
%   \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2\gamma}{2(\gamma-1)},
% \end{align}
% or $\nu< 0.1\epsilon\Omega_k$ in our disk model,
% consistent with numerical results.

% \begin{figure}
%   \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
%   \caption{Eigenvalues in for the
%     VSI in a nearly vertically isothermal disk
%     ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
%     evolved adiabatically with $\gamma=2$. The perturbation radial
%     wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
%   }
% \end{figure}

% \begin{figure}
%   \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
%   \caption{Fundamental VSI mode in a nearly vertically
%     isothermal disk ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
%     evolved adiabatically
%     with $\gamma=2$. This eigenfunction 
%     corresponds to the top-left eigenvalue displayed in 
%     Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
%     The  real (imaginary) part of $\delta v_z$ are shown as solid
%     (dotted) lines. 
%     \label{lowfreq_eigenfunc_adia}
%   }
% \end{figure}

% Although a positive vertical entropy gradient is stabilizing, we note
% that $N_z^2\propto z^2$  while vertical shear $rd\Omega^2/dz\propto
% z$, so that there is always a region sufficiently close to the
% midplane in which vertical shear dominates over bouyancy and the VSI
% can operate. However, the growth rates are small and the instability
% only affects the regions close the midplane. These modes may not be
% important in practice.

% \subsection{Other adiabatic indices}
% To see the transition from neutrally-bouyant to stably stratified cases above, 
% we repeated the previous calculation with $\gamma\in[1,2]$. 
% We restore the vertical domain to $\zmax=10\epsilon r$. 
% Fig. \ref{adia_growth_num} shows the eigenfrequencies of the
% fundamental VSI mode.

% %To select the appropriate
% %eigenvalues we vary the parameters slowly away from a reference solution.   
% % The numerically-determined eigenfrequencies are qualitatively
% % consistent with the discussion in \S\ref{adia_compare_gam1}.
% There is good agreement between Fig. \ref{adia_growth_num} and
% Fig. \ref{adia_growth} for $\khat=k_xH_\mathrm{iso}\gtrsim1$. There is a
% mismatch for smaller wavenumbers, especially in the growth
% rates. Actual growth rates decrease slower as $\khat\to0$
% than that in Fig. \ref{adia_growth}. Disagreement at small
% $\khat$ is not unexpected because the low-frequency
% approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
% $|\omega|\sim\Omega_k$ for small $\khat$.

% Given the number of approximations made in \S\ref{adia_compare_gam1},
% Fig. \ref{adia_growth_num} compares well with 
% Fig. \ref{adia_growth}. The main qualitative features of the full
% numerical solution in Fig. \ref{adia_growth_num} are captured by
% Fig. \ref{adia_growth}: stabilization for 
% $\khat\to0$ and $\khat\to\infty$,
% stabilization for increasing $\gamma$ (but ineffective at small $\khat$), 
% and the shift of the most unstable wavenumber to smaller values as
% $\gamma$ is increased.

% \begin{figure}
%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_real1}
%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag1}
%   \caption{Real frequency (top) and growth rate 
%     (bottom) of the fundamental VSI in a nearly vertically   
%     isothermal disks ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
%     subject to adiabatic perturbations with different values of $\gamma$. 
%     The eigenfrequencies are calculated by numerically
%     solving the full linear eigenvalue problem. This plot is to
%     be compared with Fig. \ref{adia_growth}. \label{adia_growth_num}}   
% \end{figure}
